---
title: "Week 04"
author: "[Callie Freitag](mailto:freitagc@uw.edu)"
date: '`r format(Sys.time(), "%Y-%m-%d %H:%M")`'
output: 
   bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
urlcolor: blue
bibliography: bibhw3.bib
biblio-style: "apalike"
link-citations: true
---

---
<!--these following settings control width and also make a DIV tag for black outlines for figures-->
<style type="text/css">
.main-container {
  max-width: 2400px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
.outline {
   background-color: #000000;
   padding:1px;
   display: inline-block;
}
</style>

<!-- for a wide presentation -->
<style type="text/css">
.main-container {
  max-width: 2400px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

<!-- a DIV for black outline, good for figures-->
<style>
.outline {
   background-color: #000000;
   padding:1px;
   display: inline-block;
}
</style>


```{r setup, message=FALSE, warning=FALSE}
# key setups
library(sqldf) # to use sql syntax with data frames
library(knitr) # knitr for kable tables
library(kableExtra) # pretty tables
library(sf) # simple features (GIS)
library(leaflet) # nice maps
library(tools) # md5sum
library(tidyverse)
library(readstata13) 
library(captioner)
library(viridis)
#install.packages("bookdown")

#caption setup
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")

# for having DIV tags
knitr::knit_hooks$set(class = function(before, options, envir) {
  if(before){
    sprintf("<div class = '%s'>", options$class)
  }else{
    "</div>"
  }
})

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```



```{r makedata, echo=FALSE, warning=FALSE, message=FALSE}
gradpain <- c(rep(0,11), rep(1,1), rep(2,2), rep(3,6), rep(4,8), rep(5,10),
 rep(6,8), rep(7,30), rep(8,10), rep(9,6), rep(10,2))

barplot(table(gradpain), las=1, ylab="Frequency", xlab="Pain Scale",
 main="How painful is graduate school?")

students <- length(gradpain)
meangradpain <- mean(gradpain)
```

# Describing the Sample
## How Many Graduate Students Are In the Sample? 
There are `r students` students in the sample.

## What is the sample mean? 
The mean of this sample is `r meangradpain`. 


# Boxes 1 & 2: Sampling the Sample 
```{r boxes12, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
#Create a funciton with arguments c(d.vec, n) that samples d.vec with replacement to produce a subsample size n. Return a list that contains the size and mean of the sample.
f1 <- function(d.vec, n) {
  s <- sample(d.vec, size = n, replace = TRUE)
  size <- length(s)
  mean <- mean(s)
  return(print(c(size, mean)))
}

# #Use set.seed(7), then run your function using d.vec=gradpain and n=length(gradpain). Do this 100 times using a loop, storing the returned means. 
set.seed(7) 

#initialize a vector to store results
bootstrapmeans <- NULL

#bootstrap 100 times
for (i in seq(1, 100, 1)){
 #returns both values (length, mean) from function
 f1out <- f1(gradpain, length(gradpain))  
 #returns the mean
 f1mean <- f1out[2] 
 #store results 
 bootstrapmeans <- c(bootstrapmeans, f1mean)
}

meanmean <- mean(bootstrapmeans)
sdmean <- sd(bootstrapmeans)
```

## What is the mean of these 100 means? 
The mean of these 100 means is `r meanmean`.

## What is the standard deviation of these 100 means? 
The standard deviation of these 100 means is `r sdmean`. 

# Box 3: Bootstrapping Results
## What does your function return for the mean of means?
## What does your function return for the standard deviation of means?
## What does your function return for the 95% confidence interval around the mean? 

For these three questions, see the R code below. I couldn't figure out how to nest the function in the loop. I'll keep trying but I'm submitting this now. 

```{r box3, echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
# Repeat the exercise in Box 2, this time with a third argument num.samples (=100)
set.seed(7) 

#The function should return a list that contains the size of each sample, the number of samples, a vector of length num.samples with the mean of each sample, the mean of the means, the sd of the means, and the 95% confidence interval around the mean. 


f2 <- function(d.vec, n, num.samples) {
  sizes <- NULL
  means <- NULL
  for (i in seq(1, num.samples, 1)){
    #sample with replacement 
    sizesmeans <- f1(d.vec, n)
    s <- sizesmeans[1]
    m <- sizesmeans[2]
    sizes <- c(sizes, s)
    means <- c(means, m)
  }
  mm <- mean(means)
  sdm <- sd(means)
  cim <- confint(means)
  obj <- list(sizes, num.samples, means, mm, sdm, cim)
  return(obj)
}

#f2out <- f2(gradpain, length(gradpain), 100)

```




# Source Code 
Source code for this document:

[1825550_HW_4.Rmd](1825550_HW_4.Rmd)

