---
title: "CSDE 502 Winter 2021, Assignment 9"
author: "[1825550, Callie Freitag](mailto:freitagc@uw.edu)"
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue 
---

```{r}
library(captioner)
require(readstata13)
require(tidyverse)

figure_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")
```

___Explanation___:
This assignment is intended to give you more practice delving into the Add Health data set and in manipulating additional variables. 

___Instructions___: 

1. Make sure your Rmd file has no local file system dependencies (i.e., anyone should be able to recreate the output HTML using only the Rmd source file).
1. Make a copy of this Rmd file and add answers below each question. The code that generated the answers should be included, as well as the complete source code for the document.
1. Change the YAML header above to identify yourself and include contact information.
1. For any tables or figures, include captions and cross-references and any other document automation methods as necessary.
1. Make sure your output HTML file looks appealing to the reader.
1. Upload the final Rmd to your github repository.
1. Download [`assn_id.txt`](http://staff.washington.edu/phurvitz/csde502_winter_2021/assignments/assn_id.txt) and include the URL to your Rmd file on github.com.
1. Create a zip file from your copy of `assn_id.txt` and upload the zip file to the Canvas site for Assignment 9. ___The zip file should contain only the text file. Do not include any additional files in the zip file--everything should be able to run from the file you uploaded to github.com. Please use zip format and not 7z or any other compression/archive format.___


#
__Using the full household roster (you'll need to go back the full raw data source, [21600-0001-Data.dta](http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip)), create the following variables for each respondent. Document any decisions that you make regarding missing values, definitions, etc. in your narrative as well as in the R code.  Include a frequency tabulation and a histogram of each result.__

Note to Phil: As I mentioned in my email, I couldn't get the code below to download and read in the file using the temporary directory system. So, I saved a copy locally and completed the exercise that way instead. 

```{r, echo=T, results='hide'}
##SELF-CONTAINED METHOD -- works to download, but can't read in file
# temp <- Sys.getenv("TMPDIR")
# 
# dtaurl <- "http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip"
# 
# zip <- file.path(Sys.getenv("TMPDIR"), basename(dtaurl))
# 
# # dta file in $temp
# dtafile <- tools::file_path_sans_ext(zip)
# 
# # check if the dta file exists
# if(!file.exists(dtafile)){
#     # if the dta file doesn't exist, check for the zip file
#     if(!file.exists(zip)){ # check if the zip file exists, download if necessary
#         curl::curl_download(url = dtaurl, destfile = zip) 
#         }
#     unzip(zipfile = zip, exdir = Sys.getenv("TEMP")) # unzip the downloaded zip file
# }
# 
# # if the data set has not been read, read it in 
# if(!exists("ahcomplete")){
#     ahcomplete <- read.dta13(dtafile)
# }
# 
# # lowercase column names
# colnames(ahcomplete) %<>% str_to_lower()


### read in data from local source
data <- haven::read_dta("/Users/freitagc/Downloads/21600-0001-Data.dta")

### check attributes of hh roster vars 
attributes(data$H1HR3A)
attributes(data$H1HR3B)
attributes(data$H1HR3C)
attributes(data$H1HR3D)
attributes(data$H1HR3E)
attributes(data$H1HR3F)
attributes(data$H1HR3G)
attributes(data$H1HR3H)
attributes(data$H1HR3I)
attributes(data$H1HR3J) #10th hh member
attributes(data$H1HR3T) #20th hh member

## check distribution of hh members
summary(data$H1HR3C) #97 is 3rd quartile
summary(data$H1HR3D) #97 is 3rd quartile
summary(data$H1HR3E) #97 is 1st quartile
summary(data$H1HR3J) #97 is 1st quartile 
summary(data$H1HR3T) #96 is min -- refused/skipped
```



Then, I got stuck in the recoding variables process. I wanted to recode the household roster variables and set the skipped values to missing. I tried to follow the code from the course website and last week's answer key (using mutate_at for the first time! Thanks for that opportunity!). But, this code ends up setting EVERYTHING to NA. What's going on? 

```{r, echo=T}
#recode vars 
hh <- data %>% select(contains("H1HR3"), AID) %>%
            mutate_at(.vars = vars(-AID),
                         list(
                             ~case_when(
                                 . == 96 ~ NA,
                                 . == 97 ~ NA,
                                 . == 98 ~ NA
                             )
                         ))

head(hh)
```


##
__Total number in household__

##
__Number of sisters__


##
__Number of brothers__


##
__Total number of siblings__


#
__What proportion of students live with two biological parents? Include the analysis in your R code.__


#
__Calculate the number of household members that are NOT biological mother, biological father, full brother or full sister. Create a contingency table and histogram for this variable.__

## Source code
```{r comment=''}
cat(readLines(con = "csde502_2021_assignment09.Rmd"), sep = '\n')
```
